\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-08 paper/draft.tex
Corrections

At a high level, \sys takes as input a POSIX shell script like the one in Fig.~\ref{fig:example} and outputs a new script, a fragment of which is shown in Fig.~\ref{fig:example2}.
% concurrent version
% To expose and exploit data parallelism, % in a given shell script,
% %% parallel execution,
% \sys analyzes individual program fragments, addressing the challenges mentioned earlier.
% 
% \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model~\sx{ir} by identifying and transforming special regions termed \emph{parallelizable regions} explained below. % nv idea: let's call them may-par regions
% It then runs a series of iterative analyses and transformations within these regions aimed at pushing parallelism as far as possible without risking breaking semantics.
% Finally, the compiler converts the DFG back into a shell script, replacing nodes with instantiated commands or primitives taken from its runtime library~\sx{optimizer}.
% An overview of \sys is shown in Fig.~\ref{fig:overview} and described below.
% First, \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model % ~\sx{dataflow-regions}
%   by identifying and transforming special regions termed \emph{parallelizable regions} and explained below. % nv idea: let's call them may-par regions
% \kk{eurosys: I suggest to replace the above sentence with the one below. There are a couple reasons. First I don't like the term source-to-source compiler because this sentence only describes the shell-graph compilation pass. Second, I would like to have some intuition about what there regions are (rather than deferring explanation for later). Third each region is tranformed to a DFG and not the whole script.}
\sys first identifies sections of the script that lack synchronization and scheduling constraints and converts them to dataflow graphs (DFGs).
It then performs a series of DFG transformations aimed at exposing parallelism without breaking semantics. % ~\sx{ir:transformations}.
Finally, \sys converts these DFGs back to a shell script augmented with \sys-provided commands. % ~\sx{optimizer}.
The script is handed off to the user's original shell interpreter for execution.
\sys addresses the aforementioned challenges~\sx{bg:challenges} as below.

@@

\heading{Scripts}
To maintain sequential semantics, \sys first analyzes a script to identify \emph{parallelizable regions} that contain commands that are candidates for parallelization.
This analysis is guided by the script structure: % to identify parallelization candidates:
  some constructs (\eg \ttt{&}, \ttt{|}) expose task parallelism; others (\eg \ttt{&&}, \ttt{||}) enforce synchronization.
% \sys first needs to understand the POSIX constructs used to compose larger programs, indicating different levels of candidacy for pararllelization.
%% To assign a candidacy for parallelization, \sys breaks POSIX constructs into a few broad classes:
%%  (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
%%  (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
%%  (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
%%  (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.
% Since the shell representation only exposes coarse grained parallelization opportunities, 
\sys then converts each parallelizable region to a \emph{dataflow graph} (DFG), 
%% Provided information commands and POSIX constructs, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
%% To achieve this, it converts sections of the script that are amenable to parallelization into a data-flow graph (DFG) representation.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
  a flexible representation that enables a series of local transformations to expose data parallelism, converting the graph into its parallel equivalent.
%% , in which DFGs indicate the parallelizable regions.
Further transformations compile the DFG back to a shell script that uses POSIX constructs to guide parallelism explicitly while aiming at preserving the semantics of the sequential program.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/pash_dfg.pdf}}
  \vspace{-25pt}
  \caption{
    \textbf{From Shell script AST to DFGs.}
    The AST on the left has two dataflow regions, each not extending beyond \ttiny{&&}~\cf{dataflow-regions}.
  }
  \vspace{-15pt}
  \label{fig:ast-dfg}
\end{figure}

@@

\heading{Auxiliary Transformations}
%
\sys also performs a set of auxiliary transformations $t_{1-3}$that are
depicted in \Cref{fig:auxiliary-transformations}. If a node has many
inputs, $t_1$ concatenates these inputs by inserting a \ttt{cat} node
to enable the parallelization transformations. In cases where a
parallelizable node has one input and is not preceded by a
concatenation, $t_2$ inserts a \ttt{cat} node that is preceded by its
inverse \ttt{split}, so that the concatenation can be commuted with
the node. Transformation $t_3$ inserts a relay node that
performs the identity transformation. Relay nodes can be useful for
monitoring and debugging, as well as for performance
improvements~\sx{optimizer}.

@@

\heading{Highlights}
\sys accelerates ($1.92-61.1\times$) all scripts.\footnote{
  \nv{But this is not true; ``the vast majority''? maybe put 51/54 scripts?}.
}
For some of our benchmarks~\sx{ours}, we decouple the performance benefits stemming from \sys's compiler alone from the ones stemming from \sys's runtime primitives:
  the compiler's average speedup seen is $9.3\times$, jumping to $13.2\times$ when runtime optimizations are used.
In a different set of $30$ benchmarks parallelized $16\times$\footnote{
  \nv{Immediate question: why use $64\times$ in one and $16\times$ in another?}
}
 see an average speedup of $6\times$, depending on the class of piecewise commands~\sx{unix50}.
Two complex scripts with 15~\sx{macro1} and 34~\sx{macro2} commands in multiple languages see $2.44\times$ and $12.7\times$ speedup on $16\times$-parallel configuration.
\sys outperforms \ttt{sort --parallel} by a significant margin, and offers small performance benefits over GNU \ttt{parallel}, which returns incorrect results in one configuration~\sx{micro}.
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
% \sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
% In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

The vast majority of programs ($>50$, with $>200$ commands) require no effort to parallelize other than invoking \sys;
  only 6 commands ($<3\%$) needed a single-record annotation.
% 
In terms of correctness, \sys's results over multi-GB  inputs are identical to the sequential for all benchmarks.
Scripts feature ample opportunities for breaking semantics, which \sys avoids.
%% In one case, \sys avoids transforming a cycle (\ie a fragment whose input and output identifiers are the same).
%in another, it preserves the semantics that another parallelization command breaks.

>>>>

At a high level, \sys takes as input a POSIX shell script like the one in Fig.~\ref{fig:example} and \todo{a parameter} \ttt{--width} and outputs a new script, a fragment of which is shown in Fig.~\ref{fig:example2}.
% concurrent version
% To expose and exploit data parallelism, % in a given shell script,
% %% parallel execution,
% \sys analyzes individual program fragments, addressing the challenges mentioned earlier.
% 
% \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model~\sx{ir} by identifying and transforming special regions termed \emph{parallelizable regions} explained below. % nv idea: let's call them may-par regions
% It then runs a series of iterative analyses and transformations within these regions aimed at pushing parallelism as far as possible without risking breaking semantics.
% Finally, the compiler converts the DFG back into a shell script, replacing nodes with instantiated commands or primitives taken from its runtime library~\sx{optimizer}.
% An overview of \sys is shown in Fig.~\ref{fig:overview} and described below.
% First, \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model % ~\sx{dataflow-regions}
%   by identifying and transforming special regions termed \emph{parallelizable regions} and explained below. % nv idea: let's call them may-par regions
% \kk{eurosys: I suggest to replace the above sentence with the one below. There are a couple reasons. First I don't like the term source-to-source compiler because this sentence only describes the shell-graph compilation pass. Second, I would like to have some intuition about what there regions are (rather than deferring explanation for later). Third each region is tranformed to a DFG and not the whole script.}
\sys first identifies sections of the script that lack synchronization and scheduling constraints and converts them to dataflow graphs (DFGs).
It then performs a series of DFG transformations aimed at exposing parallelism without breaking semantics, \todo{up to the desired graph} \ttt{width}.
. % ~\sx{ir:transformations}.
Finally, \sys converts these DFGs back to a shell script augmented with \sys-provided commands. % ~\sx{optimizer}.
The script is handed off to the user's original shell interpreter for execution.
\sys addresses the aforementioned challenges~\sx{bg:challenges} as below.

@@

\heading{Scripts}
To maintain sequential semantics, \sys first analyzes a script to identify \emph{parallelizable regions} containing commands that are candidates for parallelization.
This analysis is guided by the script structure: % to identify parallelization candidates:
  some constructs expose task parallelism; others enforce synchronization.
% \sys first needs to understand the POSIX constructs used to compose larger programs, indicating different levels of candidacy for pararllelization.
%% To assign a candidacy for parallelization, \sys breaks POSIX constructs into a few broad classes:
%%  (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
%%  (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
%%  (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
%%  (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.
% Since the shell representation only exposes coarse grained parallelization opportunities, 
\sys then converts each parallelizable region to a \emph{dataflow graph} (DFG), 
%% Provided information commands and POSIX constructs, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
%% To achieve this, it converts sections of the script that are amenable to parallelization into a data-flow graph (DFG) representation.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
  a flexible representation that enables a series of local transformations to expose data parallelism, converting the graph into its parallel equivalent.
%% , in which DFGs indicate the parallelizable regions.
Further transformations compile the DFG back to a shell script that uses POSIX constructs to guide parallelism explicitly while aiming at preserving the semantics of the sequential program.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/pash_dfg.pdf}}
  \vspace{-25pt}
  \caption{
    \textbf{From Shell script AST to DFGs.}
    The AST on the left has two dataflow regions, each not extending beyond \ttiny{&&}~\cf{dataflow-regions}.
    \nv{DFG2 reads from f2, grep writes to f3}
  }
  \vspace{-15pt}
  \label{fig:ast-dfg}
\end{figure}

@@

\heading{Auxiliary Transformations}
%
\sys also performs a set of auxiliary transformations $t_{1-3}$that are
depicted in \Cref{fig:auxiliary-transformations}. If a node has many
inputs, $t_1$ concatenates these inputs by inserting a \ttt{cat} node
to enable the parallelization transformations. In cases where a
parallelizable node has one input and is not preceded by a
concatenation, $t_2$ inserts a \ttt{cat} node that is preceded by its
inverse \ttt{split}, so that the concatenation can be commuted with
the node. Transformation $t_3$ inserts a relay node that
performs the identity transformation. Relay nodes can be useful for
monitoring and debugging, as well as for performance
improvements~\sx{optimizer}.
The additional relay nodes inserted by the auxiliary transformations affect the resulting % (task) 
  parallelism.

@@

\heading{Highlights}
\sys accelerates ($1.92-61.1\times$) \todo{49/54} scripts, whereas for the rest 5 the parallel execution time is close to the sequential (0.89, 0.91, 0.94, 099, 1.01$\times$).
In this paragraph we summarize results for \ttt{--width=16}, but the rest of our evaluation section reports on varying \ttt{width}s.
For a subset of our benchmarks~\sx{ours}, we decouple the performance benefits stemming from \sys's DFG transformations alone from the ones stemming from its runtime primitives:
  the compiler's average speedup is \todo{$9.3\times$}, jumping to \todo{$13.2\times$} when runtime optimizations are used.
A different subset ($30$ benchmarks) see an average speedup of $6.02\times$~\sx{unix50}.
% , depending on the classed of commands involved
Two complex scripts with 15~\sx{macro1} and 34~\sx{macro2} commands in multiple languages see speedups of $2.44\times$ and $12.7\times$.
\sys outperforms \ttt{sort --parallel} by a significant margin, and offers small performance benefits over GNU \ttt{parallel}, which returns incorrect results in one configuration~\sx{micro}.
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
% \sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
% In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

Using \sys's standard library of annotations for POSIX and GNU commands~\sx{parallelizability}, the vast majority of programs ($>50$, with $>200$ commands) require no effort to parallelize other than invoking \sys;
  only 6 commands outside these classes ($<3\%$) needed a single-record annotation.

In terms of correctness, \sys's results on multi-GB inputs are always identical to the sequential.
Scripts feature ample opportunities for breaking semantics, which \sys avoids.
%% In one case, \sys avoids transforming a cycle (\ie a fragment whose input and output identifiers are the same).
%in another, it preserves the semantics that another parallelization command breaks.
